---
title: "Does the diversity of a country's music preferences say anything about its 'democraticness'?"
author: "Nadia Chan"
date: "2023-12-12"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

```{r}
# libraries ---------------------------------------------------------------

library(tidyverse)
library(ggplot2)
library(stargazer)
library(readxl)
library(countrycode)
library(gridExtra)
library(lfe)


# importing data ----------------------------------------------------------

dat = read.csv('merged_data.csv') # cleaned and merged dataset


# dichotomization ---------------------------------------------------------

dat = dat %>%
  group_by(year) %>%
  mutate(
    #internet_pen = log(internet_pen),
    youth_dep_ratio = case_when(
      youth_dep_ratio > median(youth_dep_ratio, na.rm = TRUE) ~ 1,
      youth_dep_ratio <= median(youth_dep_ratio, na.rm = TRUE) ~ 0,
      is.na(youth_dep_ratio) ~ 0
    )
  )
```


## Data Sources

- Music diversity: Number of 'popular' genres in a country-year, extrapolated from *"Weekly Top 200 songs on Spotify by country from 2021 to 2022"*: Uploaded by "Yelexa" on "Kaggle"

- Democracy Index (Economist Intelligence Unit): Ranges from 0 (least democratic) to 10 (most)

- Youth dependency ratio (World Bank): People younger than 15, measured as a percentage of the working-age population (those ages 15-64). 

Controls: 

  - GDP per capita (World Bank): Measured in current US$
  
Clusters:

  - Countries categorized into 7 Regions as defined in the World Bank Development Indicators 


## Hypothesis and Model

H: Countries with more diverse music preferences, on average, have higher democracy index scores, conditional on a youth dependency ratio above the global median.

\begin{align*}
DemIndex_{it} = & \alpha + \beta_1{NumGenres_{it}} + {\beta_2}GDPPC_{it} + \beta_3{YouthDepRatio_{it}} + \\
& {\beta_4}(NumGenres_{it}*YouthDepRatio_{it}) + {\delta}\tilde{X} + \varepsilon_{it}
\end{align*}


## Summary stats

```{r}
scatter = dat %>% ggplot() +
  aes(x=num_genres, y=dem_index, color=as.factor(region)) +
  geom_point() +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(7, 'YlOrRd'),
    name = 'Region') +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='black', alpha=.25,
              linetype = 2) +
  xlab('Number of "popular" song genres') +
  ylab('Democracy Index score') +
  #ggtitle("") +
  theme_dark()
```

```{r}
scatter
```


## Results

```{r}
dat = dat %>% subset(dem_index>4 & num_genres>150)

## lm models ----
# simple cause-effect linear regression
model1 = lm(dem_index ~ num_genres, 
            data = dat)
# interaction
model2 = lm(dem_index ~ num_genres + youth_dep_ratio + num_genres*youth_dep_ratio,
            dat)
# controlled interaction
model3 = lm(dem_index ~ num_genres + gdppc + youth_dep_ratio + num_genres*youth_dep_ratio,
            dat)

## felm models ----
# region fixed effects
model4 = felm(dem_index ~ num_genres + gdppc + youth_dep_ratio + num_genres*youth_dep_ratio | region,
              dat)
# with region clustered standard errors
model5 = felm(dem_index ~ num_genres + gdppc + youth_dep_ratio + num_genres*youth_dep_ratio | region | 0 | region,
              dat)
```

\begin{tiny}

```{r, include=T, results='asis'}
stargazer(model1, model2, model3, model4, model5,
          type = 'latex',
          title = 'Core results',
          dep.var.labels = c('Democracy Index score'),
          covariate.labels = c('Number of song genres',
                               # 'GDP per capita',
                               # 'Internet Penetration'
                               # 'Literacy rate',
                               'Youth dependency ratio',
                               'Genres * Youth dependency ratio'),
          omit.stat = c('rsq', 'ser', 'f'),
          omit = c('gdppc'),
          style = 'ajps',
          model.numbers = F,
          #model.names = F,
          add.lines = list(
            c('Controlled for confounders', 'No', 'No', 'Yes', 'Yes', 'Yes'),
            c('Region Fixed Effects', 'No', 'No', 'No', 'Yes', 'Yes'),
            c('Region-level clustered S.E.', 'No', 'No', 'No', 'No', 'Yes')
          ),
          header = F)
```

\end{tiny}


## Plots

```{r}
## prediction plot ----

N = 50 # makes 50 predictions for each condition (i.e. 50 for young, 50 for not young)

predictions = data.frame(
  
  num_genres = 
    c(seq(min(dat$num_genres), max(dat$num_genres), length.out=N),
      seq(min(dat$num_genres), max(dat$num_genres), length.out=N)),
  
  youth_dep_ratio = c(rep(0,N), rep(1,N)),
  
  gdppc = mean(dat$gdppc, na.rm=T)
  
)

fits = predict(model3,
               newdata = predictions,
               interval = "confidence") 

predictions = cbind(predictions, fits)

pred_plot = predictions %>% ggplot() +
  aes(x=num_genres, y=fit, 
      ymin = lwr, ymax = upr,
      color = factor(youth_dep_ratio),
      linetype = factor(youth_dep_ratio)
  ) + 
  geom_line() +
  geom_ribbon(alpha = 0.1) + 
  scale_color_manual(name = "",
                     values = c("tomato4", 
                                "aquamarine4"), 
                     labels = c('Below or average \n youth dependency ratio',
                                'Above average \n youth dependency ratio')) +
  scale_linetype_manual(name = "",
                        values = c(2, 1), 
                        labels = c('Below or average \n youth dependency ratio',
                                   'Above average \n youth dependency ratio')) +
  theme_light() +
  xlab('Number of "popular" song genres') +
  ylab("Predicted Democracy Index score") +
  theme_light() +
  ggtitle("Prediction Plot\n(lm model)") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position='bottom', legend.direction='horizontal')
```

```{r}
# simple vector containing binary indicators for conditional value
youth_dep_ratio = c(0, 1)

# b1 = coefficient of effect MD
# b4 = coefficient of effect (MD given YDR)

# ME of MD on DI = b1 + (b4*YDR)
effect = model5$coefficients[1] + (youth_dep_ratio*model5$coefficients[4])

# SE of MD on DI = sqrt( var(b1) + (YDR)^2var(b4) + (2)(YDR)*cov(b1,b4) )
margin_se = sqrt( 
  vcov(model5)[1,1] +
    (youth_dep_ratio^2) * (vcov(model5)[4,4]) +
    2 * (youth_dep_ratio*vcov(model5)[1,4]) 
)

me_plot = cbind(youth_dep_ratio, effect, margin_se) %>% 
  as.data.frame() %>% 
  mutate(
    CI_L = effect - (1.96*margin_se),
    CI_H = effect + (1.96*margin_se)
  ) %>% 
  ggplot() +
  aes(x=youth_dep_ratio, y=effect,
      ymin=CI_L, ymax=CI_H) +
  geom_point() +
  geom_errorbar(width = 0.2) +
  scale_x_continuous(breaks = c(0,1),
                     labels = c('Below or average \n youth dependency ratio',
                                'Above average \n youth dependency\n ratio')) +
  ggtitle('Marginal effect of music diversity on\n democracy index\n(felm+c.s.e. model)') +
  xlab("") +
  ylab("") +
  geom_hline(yintercept=0, linetype=2, color='cornflowerblue', linewidth=1) +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, fig.height=4, fig.width=8.5}
grid.arrange(
  pred_plot, me_plot,
  nrow = 1
)
```


## References

\begin{tiny}

- Yelexa. "Spotify Weekly Top 200 Songs Streaming Data." \textit{Kaggle}, July 25, 2022. https://www.kaggle.com/datasets/yelexa/spotify200.

- "Democracy Index 2021: The China Challenge." \textit{Economist Intelligence Unit}, September 28, 2023. https://www.eiu.com/n/campaigns/democracy-index-2021/.

- "Democracy Index 2022." \textit{Economist Intelligence Unit}, September 28, 2023. https://www.eiu.com/n/campaigns/democracy-index-2021/.

- "Individuals Using the Internet (Percentage of Population)." \textit{World Bank Open Data}. Accessed December 12, 2023. https://data.worldbank.org/indicator/IT.NET.USER.ZS.

- "GDP per Capita (Current US Dollars)." \textit{World Bank Open Data}. Accessed December 12, 2023. https://data.worldbank.org/indicator/NY.GDP.PCAP.CD.

\end{tiny}


## Extra: Summary Stats by region

```{r}
dat = read.csv('merged_data.csv')
```

```{r}
EastAsia = dat %>% subset(region == 'East Asia & Pacific') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='gold3', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('East Asia & Pacific') +
  theme(plot.title = element_text(hjust = 0.5))

Europe = dat %>% subset(region == 'Europe & Central Asia') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='gold3', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('Europe & Central Asia') +
  theme(plot.title = element_text(hjust = 0.5))

LatinAmerica = dat %>% subset(region == 'Latin America & Caribbean') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='maroon', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('Latin America & Caribbean') +
  theme(plot.title = element_text(hjust = 0.5))

MiddleEast = dat %>% subset(region == 'Middle East & North Africa') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='maroon', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('Middle East & North Africa') +
  theme(plot.title = element_text(hjust = 0.5))

NorthAmerica = dat %>% subset(region == 'North America') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='maroon', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('North America') +
  theme(plot.title = element_text(hjust = 0.5))

SouthAsia = dat %>% subset(region == 'South Asia') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='gold3', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('South Asia') +
  theme(plot.title = element_text(hjust = 0.5))

SubsaharanAfrica = dat %>% subset(region == 'Sub-Saharan Africa') %>% 
  ggplot() +
  aes(x=num_genres, y=dem_index) +
  geom_point() +
  geom_smooth(formula=y~x, method='lm', 
              fill='grey', color='gold3', alpha=.25,
              linetype = 2) +
  xlab('Number of genres') +
  ylab('DI score') +
  xlim(100, 450) +
  ylim(-10, 10) +
  theme_minimal() +
  ggtitle('Sub-Saharan Africa') +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
grid.arrange(EastAsia, Europe, MiddleEast, LatinAmerica,
             NorthAmerica, SouthAsia, SubsaharanAfrica,
             nrow = 4)
```


## Extra: Why median as cutoff for interaction term?

```{r}
dat$youth_dep_ratio %>% hist(main = "Histogram of Youth dependency ratio")

```




